---
title: "TS Individual Assignment"
author: "Paarth Tiwari"
date: "3/8/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
#Importing all the relevant libraries
rm(list=ls())
library('forecast')
library('tseries')
library('stats')
library('tidyverse')
```


## 1.	Loading the dataset, initial observations and plots

The dataset going to be used for this case study is the gas dataset from the forecast package. This dataset contains the Australian monthly gas production data from 1956 to 1995.

### i.	Data loading
```{r}
# Loading the dataset
data(gas)
# Creating TS object
gasTS <- ts(gas,
            start = c(1956,1),
            end = c(1995,8),
            frequency = 12)
class(gasTS)
#force(gasTS)
#View(gasTS)
str(gasTS)
```

**Observations:**  
a. gas is a time-series object with 476 observations distributed monthly from 1956 to 1995.
b. No missing data observed in the time series.

### ii.	Basic exploratory TS Plots

+ _Plot 1 - TS plot_
  
```{r}
plot(gasTS)
abline(reg = lm(gasTS~time(gasTS)))
```

**Observations:**   
Some of the impressions from the basic TS plot -  
a. From 1956 to 1969, the yearly gas production appears to be relatively uniform as the graph does not show any obvious trend in the gas production pattern.  
b. Post 1970 to 1982, a sharp upward/positive trend can be observed in the production of gas units.  
c. Post 1982, yearly production pattern is relatively uniform. However, intra-year fluctuations in the monthly productions patterns are observed indicating presence of seasonality component in the time series.  

+ _Plot 2 - Month Plot_
```{r}
monthplot(gasTS)
```

**Observations:**  
a. There appears to be an almost bell-curve like distribution in the gas production pattern with mean production being lowest in January and December with the highest mean production in July.   
b. While the overall production shows an upward trend, we can observe a dip in the production levels towards the later years of the time series.  

As we proceed further, we'll analyze these observations in detail.   

For further analysis, we'll be subsetting the data from 1970(as prescribed in the problem statment)  

```{r}
# Taking a subset of the series for analysis
gas.subset <- window(gas,start = 1970)
#force(gas.subset)
```


Let's create plots for observing the seasonality in the dataset

+ _Plot 3 - Seasonal Plot_

```{r}
ggseasonplot(gas.subset, 
             year.labels=TRUE, 
             year.labels.left=TRUE) + 
  ylab("degree") + 
  ggtitle("Seasonal plot: Gas Production Data")
```

**Observations:**  
As indicated in the earlier monthly plot, for any given year, the production peaks in July.

+ _Plot 4 - Polar Seasonal plot Year-wise_

```{r}
ggseasonplot(gas.subset, 
             polar=TRUE) + 
  ylab("degree") + 
  ggtitle("Polar seasonal plot: Gas Production Data")
```

**Observations:**  
Year-on-year production of gas units is increasing.

## 2.	Time series components of the dataset

Thus far, the given dataset exhibits the characteristics of a multiplicative series.  
We'll now decompose the time series to identify the trend, seasonality and irregular components present in the given series.  

### i. Decomposition of TS using slt() 

+ _By assuming seasonlity is uniform in each year_ - 
```{r}
# Decomposition of gas TS using stl()
gas.subset.decompose1<-stl(gas.subset, s.window='p') 
#gas.subset.decompose1
plot(gas.subset.decompose1)
```

**Observations:**  
From the scales of the above graphs, it can be interpreted that Trend and Seasonality components impact the production of gas units.  

+ _By assuming seasonlity is non-uniform_ - 

```{r}
gas.subset.decompose2<-stl(gas.subset, s.window=7) 
#gas.subset.decompose2
plot(gas.subset.decompose2)
```

**Observations:**  
Seasonality appears to be increasing over time. Still the trend component is more dominant.


### i. Decomposition of TS using decompose() 

```{r}
gas.subset.decompose<-decompose(gas.subset, type = "multiplicative")
plot(gas.subset.decompose)

```

_Checking the seasonal indices_ - 

```{r}
## Observing the seasonal indices of gas.subset.decompose
Seasonal_Ind=round(t(gas.subset.decompose$figure),2) 
colnames(Seasonal_Ind)<-c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec") 
Seasonal_Ind
```


**Observations:**  
The above indices confirm the observation that production is highest in July and lowest in January.


## 3.	Periodicity

### •	Test and train splits

```{r}
gas.subset.train <- window(gas.subset, 
                   start=c(1970,1), 
                   end=c(1993, 12), 
                   freq=12) 

gas.subset.test <- window(gas.subset, 
                  start=c(1994,1), 
                  freq=12)

#force(gas.subset.train)
#force(gas.subset.test)
```


## 4.	Stationarity

For a time series is to qualify as stationary, its mean and variance should remain constant over time. There should be no trend component i.e. series should have a steady best fit line. 

To test whether given series is stationary, we will use the following tests - 

### i.	Visual inspection and KPSS test

In the earlier plots in Section 1, we have observed that with time, the fluctuation about the mean is not constant for the given series. Therefore, the gas timeseries exhibits a non-constant variance and is not stationary.

We can also confirm this using the KPSS(Kwiatkowski–Phillips–Schmidt–Shin) test to determine if the series is stationary.  


KPSS test for stationarity is defined as -   
H0: Series is stationary  
H1: Series is non-stationary  

```{r}
kpss.test(gas.subset.train)
```
**Test conclusion:**
KPSS test returned p-value = 0.01 < 0.05 i.e. H0 is rejected
Therefore, it can be concluded,with 95% confidence,  that the given series is non-stationary.

### ii.	ADF test

We will now use the ADF(Augmented Dickey-Fuller) test to check the stationarity of the given series.

ADF test for stationarity is defined as - 
H0: Series is non-stationary
H1: Series is stationary

```{r}
adf.test(gas.subset.train,k=24)
```
**Test conclusion:**
ADF test returned p-value = 0.98 > 0.05 i.e. H0 is accepted
Therefore, it can be concluded,with 95% confidence, that the given series is non-stationary.  

In order to apply ARIMA forecasting models, the series needs to be made stationary.  

### iii.	Steps to de-seasonalize the series

In order to make the variance constant, we can use logarithmic or trignometric transformations - 

Using log10 transformation:

```{r}
plot(log10(gas.subset.train))
```

**Observations:**  
Variance appears nearly constant.  

In order to achieve a constant mean, we can use differencing - 

```{r}
plot(diff(log10(gas.subset.train)))
```

**Observations:**  
Mean and variance appear nearly constant. 


Let's test the transformed series for stationarity:
```{r}
# Transformed series 
tran.gas.subset.train <- diff(log10(gas.subset.train))

# KPSS test
kpss.test(tran.gas.subset.train)
```
**Test conclusion:**
KPSS test returned p-value = 0.1 > 0.05 i.e. H0 is accpeted
Therefore, it can be concluded,with 95% confidence,  that the transformed series is stationary.

```{r}
# ADF test
adf.test(tran.gas.subset.train,k=24)

```
**Test conclusion:**
ADF test returned p-value = 0.037 < 0.05 i.e. H0 is rejected
Therefore, it can be concluded,with 95% confidence, that the transformed series is stationary.  

## 5.	Forecasting

### i.	Initial forecast for next 20 periods

### ii.	Final forecast for next 12 periods

## 6.	Accuracy of the model

